---
title: "Hydrogen_FP"
author: "A.Mehta,A.Cheekoti,A.Tayada"
output: pdf_document
---

Chunk 1: Defining global parameters.
```{r}
rm(list=ls())

N_dim = 3                         # Number of dimensions 

alpha_inc = 0.1                  # Incrememnt in alpha values
num_alphas=10                     # Total number of alphas 

num_walkers=100                    # Number of walkers
num_steps=10000                   # Number of total steps by each walker (thermalisation + sampling)
therm_steps=0.2*num_steps         # Number of thermalisation steps

del_t=0.01
therm_interval=(therm_steps)/10   # Interval size to adjust step size

```

Chunk 2: Defining functions for sampling and local energy.
```{r}
# Function to calculate norm of a vector
norm_vec <- function(x) sqrt(sum(x^2))

# Function to sample position values from (Psi^2/<Psi|Psi>)
rho <- function(x,alpha) {
  return ((norm_vec(x)^(3-N_dim))*exp(-2*alpha*(norm_vec(x))))
}

# Function to calculate local energy
elocal <- function(x,alpha) {
  return (-0.5*(alpha^2 - 2*(alpha/x)) - (1/x))
}
#Force
F_fp <-function(r,alpha){
  return(-2*alpha*r/norm_vec(r))
}
#Green's function
G_fp <-function(x,y,del_t,alpha){
  power=sum((y-x-(F_fp(x,alpha)*del_t/2))^2)
  return(exp(-power/(del_t*2))/((2*pi*del_t)^(N_dim/2)))
}
```

Chunk 3: Executing variational monte carlo with calculation of local energy and variance.
```{r}
alpha=0.5
r_total=c()
E_local_sum_all = c()
E_local_sqd_sum_all = c()

alpha_all = c()

accept_rate_all = matrix(0,nrow=num_alphas,ncol = num_walkers)

for (i in seq(num_alphas)){ #For different alphas
  
  E_local_sum_one_alpha = 0
  E_local_sqd_sum_one_alpha = 0
  alpha_all = c(alpha_all,alpha)
  
  for (j in seq(num_walkers))
    {
        r_sample=c()
        r= runif(N_dim,-5,5)            # Initial position of the walker
        accept = 0
    
        for(k in seq(num_steps))   # Generate position values for each alpha
             { 
              r_dum=r+(F_fp(r,alpha)*del_t/2)+(rnorm(N_dim)*sqrt(del_t))
              ratio = (G_fp(r_dum,r,del_t,alpha)*rho(r_dum,alpha))/(G_fp(r,r_dum,del_t,alpha)*rho(r,alpha))
      
             if (ratio>=runif(1,0,1))
                    {
                         r_sample = c(r_sample,norm_vec(r_dum))
                         r=r_dum
                         accept = accept + 1
                    } else{r_sample = c(r_sample,norm_vec(r))}
      
              if (k<=therm_steps)        #Thermalisation steps
                    {
              if (k%%therm_interval==0) #Changes in step size after uniform intervals
                    {
                        accept=max(1,accept)
                        del_t= del_t*((10/9)*accept/therm_interval)
                        accept=0
                    }
              }
         } #one walker
          
        E_local_one_walker = elocal(r_sample[(therm_steps+1):num_steps],alpha)    # array of one particular walker, energies for all x
        E_local_sum_one_alpha = E_local_sum_one_alpha + sum(E_local_one_walker)   # Adding all local energy for all x, all walkers, one alpha
    
        E_local_sqd_sum_one_alpha = E_local_sqd_sum_one_alpha + sum(E_local_one_walker * E_local_one_walker)
    
        accept_rate_all[i,j]=accept/(num_steps-therm_steps)
    
    }#one alpha  
  
  E_local_sum_all= c(E_local_sum_all,E_local_sum_one_alpha)  #array containing elocal for different alphas
  E_local_sqd_sum_all = c(E_local_sqd_sum_all,E_local_sqd_sum_one_alpha) #array containing elocal^2 for diff. alphas
  cat(i, "of", num_alphas, "alphas calculated.","\n")
  alpha = alpha + alpha_inc
}

E_local_var = E_local_sqd_sum_all/(num_walkers*(num_steps-therm_steps)) - (E_local_sum_all/(num_walkers*(num_steps-therm_steps)))^2

E_local_all = E_local_sum_all/(num_walkers*(num_steps-therm_steps))
std_dev = sqrt(E_local_var)
error = std_dev/sqrt(num_walkers*(num_steps-therm_steps))

```

Chunk 4: Main plots of our analysis.
```{r}
plot(alpha_all,rowMeans(accept_rate_all),xlab="alpha",ylab="Acceptance rate", main = "Acceptance rate vs alpha")
plot(alpha_all,E_local_var,xlab="alpha",ylab="Variance", main = "Variance vs alpha")

plot(alpha_all,E_local_all,xlab="alpha", ylab="Local Energy",ylim=range(c(E_local_all-error, E_local_all+error)), main = "Local energy vs alpha")
arrows(alpha_all, E_local_all-error, alpha_all, E_local_all+error , length=0.05, angle=90, code=3)

```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
