---
title: "Helium_FP"
output: html_notebook
author: A. Mehta, A. Cheekoti, A. Tayade
---

Chunk 1: Defining global parameters.
```{r}
rm(list=ls())

alpha_inc = 0.05                  # Incrememnt in alpha values
num_alphas=10                     # Total number of alphas 

num_walkers=100                    # Number of walkers
num_steps=10000                   # Number of total steps by each walker (thermalisation + sampling)
therm_steps=0.2*num_steps         # Number of thermalisation steps

del_t=0.1                          
therm_interval=(therm_steps)/10   # Interval size to adjust step size
```

Chunk 2: Defining functions for norm of vector, sampling and local energy.
```{r}
# Function to calculate norm of a vector
norm_vec <- function(r) {return(sqrt(sum(r^2)))}

#Funtion to calculate the distance between two vecors/points
r12<- function(r1,r2) {return(norm_vec(r1-r2))}

# Function to sample position values from (Psi^2/<Psi|Psi>)
rho <- function(r1,r2,alpha) {
  value_r12=r12(r1,r2)
  return (exp(2*(-2*norm_vec(r1) - 2*norm_vec(r2) + (value_r12/(2 + 2*alpha*value_r12 )) )))
}

# Function to calculate local energy
elocal <- function(r1,r2,alpha) {
  value_r12=r12(r1,r2)
  denom= 1 + alpha*value_r12
  elocal = -4 - (1/(value_r12*(denom^3))) - (1/(4*(denom^4))) + (1/value_r12) +(((r1-r2)%*%((r1/norm_vec(r1))-(r2/norm_vec(r2))))/ (value_r12*denom^2))
  return (elocal)
  }
#Force
F_fp <-function(r1,r2,alpha) {
  value_r12=r12(r1,r2)
        denom= 1 + alpha*value_r12
    forc = 2*( -(2*r1/norm_vec(r1)) - ((0.5*alpha*(r1-r2))/(denom)^2) + (0.5*(r1-r2)/(denom*value_r12))  )
  return(forc)
}
#Green's function
G_fp <-function(x1,x2,y1,y2,del_t,alpha){
  power=sum((c(y1,y2)-c(x1,x2)-(c(F_fp(x1,x2,alpha),F_fp(x2,x1,alpha))*del_t/2))^2)
   return( exp(-power/(del_t*2)) / ((2*pi*del_t)^(6/2)) )         # 6 - Number of degrees of freedom
}  
```

Chunk 3: Executing variational monte carlo with calculation of local energy and variance.
```{r}
alpha = 0.05                       # Initial value of alpha (variable parameter)

E_local_sum_all = c()
E_local_sqd_sum_all = c()

alpha_all=c()

r1_sample_all=c()
r2_sample_all=c()

#matrix to contain the acceptance rate of each walker for every alpha
accept_rate_all = matrix(0,nrow=num_alphas,ncol = num_walkers) 

for (i in seq(num_alphas)){ #For different alphas
  
  alpha_all = c(alpha_all,alpha)
  E_local_sum_one_alpha = 0
  E_local_sqd_sum_one_alpha = 0
  
  for (j in seq(num_walkers)){
   
      r1_sample=c()
      r2_sample=c()
      r1=runif(3,-2,+2)      #initial positions of the walkers
      r2=runif(3,-2,+2)
      accept=0
      
      for(k in seq(num_steps)){    #Generate r_samples for each walker
        
            r1_dum= r1 + (F_fp(r1,r2,alpha)*del_t/2) + (rnorm(3)*sqrt(del_t))
            r2_dum= r2 + (F_fp(r2,r1,alpha)*del_t/2) + (rnorm(3)*sqrt(del_t))
            
            ratio = ( G_fp(r1_dum,r2_dum,r1,r2,del_t,alpha) * rho(r1_dum,r2_dum,alpha) ) / ( G_fp(r1,r2,r1_dum,r2_dum,del_t,alpha) * rho(r1,r2,alpha) )
            
            if (ratio>=runif(1,0,1)){
                  
                  r1_sample = rbind(r1_sample,r1_dum)
                  r2_sample = rbind(r2_sample,r2_dum)
                  r1=r1_dum
                  r2=r2_dum
                  accept = accept + 1
            } else{r1_sample = rbind(r1_sample,r1)
                  r2_sample = rbind(r2_sample,r2)}
            
            #Thermalisation steps &&
            #Changes in step size after uniform intervals
            if (k<=therm_steps && k%%therm_interval==0){
                del_t= del_t*((10/9)*(accept/therm_interval))
                accept=0
            }
        
        } #one walker
      
      E_local_one_walker=c()
      for (l in seq(num_steps-therm_steps)){
          E_local_one_walker          =c(E_local_one_walker,elocal(r1_sample[(therm_steps+l),],r2_sample[(therm_steps+l),],alpha))         # array of one particular walker, energies for all x        
          }
      E_local_sum_one_alpha = E_local_sum_one_alpha + sum(E_local_one_walker)     #Adding all local energy for all x, all walkers, one aplha
      
      E_local_sqd_sum_one_alpha = E_local_sqd_sum_one_alpha + sum(E_local_one_walker^2)
      
      #old approach
      accept_rate_all[i,j]=accept/(num_steps-therm_steps)
    
    }#one alpha  
  
  #array containing elocal for different alphas
  E_local_sum_all= c(E_local_sum_all,E_local_sum_one_alpha)  
  #array containing elocal^2 for diff. alphas
  E_local_sqd_sum_all = c(E_local_sqd_sum_all,E_local_sqd_sum_one_alpha) 
  
  alpha = alpha + alpha_inc
  cat(i, "of", num_alphas, "alphas calculated.","\n")
}

E_local_var = E_local_sqd_sum_all/(num_walkers*(num_steps-therm_steps)) - (E_local_sum_all/(num_walkers*(num_steps-therm_steps)))^2

E_local_all = E_local_sum_all/(num_walkers*(num_steps-therm_steps))
std_dev = sqrt(E_local_var)
error = std_dev/sqrt(num_walkers*(num_steps-therm_steps))

```

Chunk 4: Main plots of our analysis.
```{r}
plot(alpha_all,rowMeans(accept_rate_all),xlab="alpha",ylab="Acceptance rate", main = "Acceptance rate vs alpha")
plot(alpha_all,E_local_var,xlab="alpha",ylab="Variance", main = "Variance vs alpha")

plot(alpha_all,E_local_all,xlab="alpha", ylab="Local Energy",ylim=range(c(E_local_all-error, E_local_all+error)), main = "Local energy vs alpha")
arrows(alpha_all, E_local_all-error, alpha_all, E_local_all+error , length=0.05, angle=90, code=3)
```

Chunk 5: Additional plots for reference.
```{r}
#hist(x_sample)
plot(alpha_all,rowMeans(E_trial_all),xlab="alpha",ylab="Mean Local Energy")
plot(alpha_all,rowMeans(E_trial_all),xlab="alpha",ylab="Mean Local Energy",ylim=range(c(E_local_all-error, E_local_all+error)))
arrows(alpha_all, E_local_all-error, alpha_all, E_local_all+error , length=0.05, angle=90, code=3)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
